taxize screencast
==========

```{r setup, include=FALSE, echo=FALSE}
knitr::opts_chunk$set(
  message = FALSE,
  warning = FALSE
)
```


## Intro

> Hi, I'm Scott Chamberlain

> In this screencast, I'll do a brief intoduction to the taxize package - a taxonomic toolbelt for R

> You can install taxize from [CRAN](http://cran.r-project.org/web/packages/taxize/index.html), or install the [development version from GitHub](https://github.com/ropensci/taxize).

> taxize interacts with about 20 data sources

```{r child="data_sources.md"}
```

> More will be added in the future

> Taxonomic names are something many biologists have to deal with

> I'll walk you through a variety of things you can do with taxize

## Installation and load

```{r installation, eval=FALSE}
install.packages("taxize")
```

```{r}
library(taxize)
```

## Taxonomic identifiers

```{r identifiers, eval=TRUE}
get_ids(names="Chironomus riparius", db = 'ncbi')
get_ids(names=c("Chironomus riparius", "Poa annua"), db = c('ncbi','itis','col'))
```

## Hierarchies

> and now taxonomic hieararchies


```{r hierarchies, eval=TRUE}
classification("Chironomus riparius", db = 'ncbi')
classification(get_uid("Chironomus riparius"))
# out <- get_ids(names="Poa annua", db = c('ncbi','itis','col','eol','tropicos'))
# classification(out)
```

## Check name spelling

> Check name spelling

```{r spelling, eval=TRUE}
mynames <- c("Helianthus annus", "Poa anua", "Pinus contort")
out <- tnrs(query = mynames, source="iPlant_TNRS")
out[,c(1:5)]
```

## Synonyms

> Moving on to synonyms

> The names you currently have in your dataset may be synonyms of real names, and you need to update to the accepted names

```{r synonyms, eval=TRUE}
synonyms(c("Poa annua",'Pinus contorta','Puma concolor'), db="itis")
```

## Downstream

> Sometimes you just wann know "what species are in this genus!!!"

```{r downstream, eval=TRUE}
col_downstream(name = "Apis", downto="Species")[[1]]
```

## Common from scientific and vice versa

> You may often need to get common names from scientific names, and vice versa. For example, maybe you need common names for a publication.

```{r}
sci2comm("Quercus lobata", db="itis",)
comm2sci("valley oak", db="itis")
```

## end

> That's it! Thanks for watching...
