---
layout: post
title: taxize
---

```{r, echo=FALSE}
knitr::opts_chunk$set(
	comment = "#>", 
	collapse = TRUE,
	warning = FALSE, 
	message = FALSE
)
```

Here's the taxize screencast!

Follow along below in your favorite R client.

<iframe src="//player.vimeo.com/video/92883063" width="600" height="381" frameborder="0" webkitallowfullscreen mozallowfullscreen allowfullscreen></iframe>


taxize screencast
==========


## Intro

> Hi, I'm Scott Chamberlain

> In this screencast, I'll do a brief intoduction to the taxize package - a taxonomic toolbelt for R

> You can install taxize from [CRAN](http://cran.r-project.org/web/packages/taxize/index.html), or install the [development version from GitHub](https://github.com/ropensci/taxize).

> taxize interacts with about 20 data sources


<table>
<colgroup>
<col style="text-align:left;"/>
<col style="text-align:left;"/>
<col style="text-align:left;"/>
<col style="text-align:left;"/>
<col style="text-align:left;"/>
</colgroup>

<thead>
<tr>
  <th style="text-align:left;">Souce</th>
  <th style="text-align:left;">Function prefix</th>
  <th style="text-align:left;">API Docs</th>
  <th style="text-align:left;">API key</th>
  <th style="text-align:left;"><em>soap branch only</em></th>
</tr>
</thead>

<tbody>
<tr>
  <td style="text-align:left;">Encylopedia of Life</td>
  <td style="text-align:left;"><code>eol</code></td>
  <td style="text-align:left;"><a href="http://www.eol.org/api/">link</a></td>
  <td style="text-align:left;"><a href="http://eol.org/users/register">link</a></td>
  <td style="text-align:left;">false</td>
</tr>
<tr>
  <td style="text-align:left;">Taxonomic Name Resolution Service</td>
  <td style="text-align:left;"><code>tnrs</code></td>
  <td style="text-align:left;"><a href="http://api.phylotastic.org/tnrs">link</a></td>
  <td style="text-align:left;">none</td>
  <td style="text-align:left;">false</td>
</tr>
<tr>
  <td style="text-align:left;">Integrated Taxonomic Information Service</td>
  <td style="text-align:left;"><code>itis</code></td>
  <td style="text-align:left;"><a href="http://www.itis.gov/ws_description.html">link</a></td>
  <td style="text-align:left;">none</td>
  <td style="text-align:left;">false</td>
</tr>
<tr>
  <td style="text-align:left;">Phylomatic</td>
  <td style="text-align:left;"><code>phylomatic</code></td>
  <td style="text-align:left;"><a href="http://www.phylodiversity.net/phylomatic/phylomatic_api.html">link</a></td>
  <td style="text-align:left;">none</td>
  <td style="text-align:left;">false</td>
</tr>
<tr>
  <td style="text-align:left;">uBio</td>
  <td style="text-align:left;"><code>ubio</code></td>
  <td style="text-align:left;"><a href="http://www.ubio.org/index.php?pagename=xml_services">link</a></td>
  <td style="text-align:left;"><a href="http://www.ubio.org/index.php?pagename=form">link</a></td>
  <td style="text-align:left;">false</td>
</tr>
<tr>
  <td style="text-align:left;">Global Names Resolver</td>
  <td style="text-align:left;"><code>gnr</code></td>
  <td style="text-align:left;"><a href="http://resolver.globalnames.org/api">link</a></td>
  <td style="text-align:left;">none</td>
  <td style="text-align:left;">false</td>
</tr>
<tr>
  <td style="text-align:left;">Global Names Index</td>
  <td style="text-align:left;"><code>gni</code></td>
  <td style="text-align:left;"><a href="https://github.com/dimus/gni/wiki/api">link</a></td>
  <td style="text-align:left;">none</td>
  <td style="text-align:left;">false</td>
</tr>
<tr>
  <td style="text-align:left;">IUCN Red List</td>
  <td style="text-align:left;"><code>iucn</code></td>
  <td style="text-align:left;"><a href="https://www.assembla.com/spaces/sis/wiki/Red_List_API?version=3">link</a></td>
  <td style="text-align:left;">none</td>
  <td style="text-align:left;">false</td>
</tr>
<tr>
  <td style="text-align:left;">Tropicos</td>
  <td style="text-align:left;"><code>tp</code></td>
  <td style="text-align:left;"><a href="http://services.tropicos.org/help">link</a></td>
  <td style="text-align:left;"><a href="http://services.tropicos.org/help?requestkey">link</a></td>
  <td style="text-align:left;">false</td>
</tr>
<tr>
  <td style="text-align:left;">Plantminer</td>
  <td style="text-align:left;"><code>plantminer</code></td>
  <td style="text-align:left;"><a href="http://www.plantminer.com/help">link</a></td>
  <td style="text-align:left;"><a href="http://www.plantminer.com/help">link</a></td>
  <td style="text-align:left;">false</td>
</tr>
<tr>
  <td style="text-align:left;">Theplantlist dot org</td>
  <td style="text-align:left;"><code>tpl</code></td>
  <td style="text-align:left;">**</td>
  <td style="text-align:left;">none</td>
  <td style="text-align:left;">false</td>
</tr>
<tr>
  <td style="text-align:left;">Catalogue of Life</td>
  <td style="text-align:left;"><code>col</code></td>
  <td style="text-align:left;"><a href="http://www.catalogueoflife.org/colwebsite/content/web-services">link</a></td>
  <td style="text-align:left;">none</td>
  <td style="text-align:left;">false</td>
</tr>
<tr>
  <td style="text-align:left;">Global Invasive Species Database</td>
  <td style="text-align:left;"><code>gisd</code></td>
  <td style="text-align:left;">***</td>
  <td style="text-align:left;">none</td>
  <td style="text-align:left;">false</td>
</tr>
<tr>
  <td style="text-align:left;">National Center for Biotechnology Information</td>
  <td style="text-align:left;"><code>ncbi</code></td>
  <td style="text-align:left;">none</td>
  <td style="text-align:left;">none</td>
  <td style="text-align:left;">false</td>
</tr>
<tr>
  <td style="text-align:left;">CANADENSYS Vascan name search API</td>
  <td style="text-align:left;"><code>vascan</code></td>
  <td style="text-align:left;"><a href="http://data.canadensys.net/vascan/api">link</a></td>
  <td style="text-align:left;">none</td>
  <td style="text-align:left;">false</td>
</tr>
<tr>
  <td style="text-align:left;">International Plant Names Index (IPNI)</td>
  <td style="text-align:left;"><code>ipni</code></td>
  <td style="text-align:left;"><a href="http://www.ipni.org/link_to_ipni.html">link</a></td>
  <td style="text-align:left;">none</td>
  <td style="text-align:left;">false</td>
</tr>
<tr>
  <td style="text-align:left;">World Register of Marine Species (WoRMS)</td>
  <td style="text-align:left;"><code>worms</code></td>
  <td style="text-align:left;"><a href="http://www.marinespecies.org/aphia.php?p=webservice">link</a></td>
  <td style="text-align:left;">none</td>
  <td style="text-align:left;"><strong>TRUE</strong></td>
</tr>
<tr>
  <td style="text-align:left;">Barcode of Life Data Systems (BOLD)</td>
  <td style="text-align:left;"><code>bold</code></td>
  <td style="text-align:left;"><a href="http://www.boldsystems.org/index.php/Resources">link</a></td>
  <td style="text-align:left;">none</td>
  <td style="text-align:left;">false</td>
</tr>
<tr>
  <td style="text-align:left;">Pan-European Species directories Infrastructure (PESI)</td>
  <td style="text-align:left;"><code>pesi</code></td>
  <td style="text-align:left;"><a href="http://www.eu-nomen.eu/portal/webservices.php">link</a></td>
  <td style="text-align:left;">none</td>
  <td style="text-align:left;"><strong>TRUE</strong></td>
</tr>
<tr>
  <td style="text-align:left;">Mycobank</td>
  <td style="text-align:left;"><code>myco</code></td>
  <td style="text-align:left;"><a href="http://www.mycobank.org/Services/Generic/Help.aspx?s=searchservice">link</a></td>
  <td style="text-align:left;">none</td>
  <td style="text-align:left;"><strong>TRUE</strong></td>
</tr>
<tr>
  <td style="text-align:left;">National Biodiversity Network (UK)</td>
  <td style="text-align:left;"><code>nbn</code></td>
  <td style="text-align:left;"><a href="https://data.nbn.org.uk/Documentation/Web_Services/Web_Services-REST/resources/restapi/rest.html">link</a></td>
  <td style="text-align:left;">none</td>
  <td style="text-align:left;">false</td>
</tr>
</tbody>
</table>

> More will be added in the future

> Taxonomic names are something many biologists have to deal with

> I'll walk you through a variety of things you can do with taxize

## Installation and load


```{r eval=FALSE}
install.packages("taxize")
```


```{r}
library("taxize")
```

## Taxonomic identifiers


```{r}
get_ids(names="Chironomus riparius", db = 'ncbi')
```

```{r}
get_ids(names=c("Chironomus riparius", "Poa annua"), db = c('ncbi','itis','col'))
```

## Hierarchies

> and now taxonomic hieararchies

```{r}
classification("Chironomus riparius", db = 'ncbi')
```

```{r}
classification(get_uid("Chironomus riparius"))
```

```{r eval=FALSE}
out <- get_ids(names="Poa annua", db = c('ncbi','itis','col','eol','tropicos'))
classification(out)
```

## Check name spelling

> Check name spelling

```{r}
mynames <- c("Helianthus annus", "Poa anua", "Pinus contort")
out <- tnrs(query = mynames, source="iPlant_TNRS")
out[,c(1:5)]
```

## Synonyms

> Moving on to synonyms

> The names you currently have in your dataset may be synonyms of real names, and you need to update to the accepted names

```{r}
synonyms(c("Poa annua",'Pinus contorta','Puma concolor'), db="itis")
```

## Downstream

> Sometimes you want know "what species are in this genus!"

```{r}
col_downstream(name = "Apis", downto="Species")[[1]]
```

## Common from scientific and vice versa

> You may often need to get common names from scientific names, and vice versa. For example, maybe you need common names for a publication.

```{r}
sci2comm("Quercus lobata", db="itis")
```

```{r}
comm2sci("valley oak", db="itis")
```

## end

> That's it! Thanks for watching...
